251 MA

### 2022-11-29 code/process.Rmd
more work on ES calculations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(viridis)
library(Replicate)
library(metafor)
library
theme_set(theme_bw())
```

# TODOs
- why are some ES's  about 2 times bigger than they should be?
- get ES from t test, from F test (using compute es) < -- what does within_one and within_two mean??
- validate df against num people?
- what about for other test statistics (r)

@@

  select(target_lastauthor_year, status, academic_year, subfield, target_year, stanford_internal, 
         open_data, open_materials, target_on_turk, replication_on_turk, 
         within_between, single_vignette, repeated_measure,
         target_N, target_test, target_raw_stat, target_pvalue, target_ES_type, target_ES_value,
         replication_N, replication_test, replication_raw_stat, replication_pvalue, replication_ES_type, replication_ES_value,
         replicated_instructor_code, replicated_report_code)

@@

We parse out values from (for now) t tests and F tests (other tests to be added later).

@@

parse_f <- function(fval, within_between){
  df_1 =str_extract(fval, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  df_2 =str_extract(fval, ",.*\\)") |> str_sub(2,-2) |> as.numeric()
  val=str_extract(fval, "=.*") |> str_sub(2,-1) |> as.numeric() |> abs()
  pval=pf(q=val, df1=df_1, df2=df_2, lower.tail=FALSE)
  d_calc=NA
  if(!is.na(df_1)&&df_1==1){
    if(within_between=="between"){
      d_calc=2*sqrt(val)/sqrt(df_2) # if there's just two groups, then F=t**2 and we can use the t-test thingy
    }
    else{
      d_calc=sqrt(val)/sqrt(df_2)
    }
  if(!is.na(df_1)){
    # we go through partial eta
    n=df_1+df_2+1 # rederive n b/c it's not always matching the n we have
    partial_eta=(val*df_1)/(val*df_1+df_2)
    d_calc=sqrt(n/(n-1)*(partial_eta/(1-partial_eta)))
  }
}
  return(data.frame("df_1"=df_1,"df_2"=df_2,"tstat"=NA, "fstat"=val, "p_calc"=pval, "d_calc"=d_calc))
}

do_blanks <- function(){
  return(data.frame("df_1"=NA,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=NA, "d_calc"=NA))
}
do_parsing=function(raw_stat, within_between){
  if (is.na(raw_stat)) {return (do_blanks())}
  if (str_sub(raw_stat,1,1)=="t"){return(parse_t(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="F"){return(parse_f(raw_stat, within_between))}
  return (do_blanks())
}
```

```{r}
test <- d_val |> 
  mutate(target=map2(target_raw_stat, within_between, do_parsing),
         rep=map2(replication_raw_stat, within_between, do_parsing)) |> 
  unnest(cols=c(target, rep), names_sep="_")

@@

Comparing calculated p vals to provided (or hand calculated) ones.

@@

test |> rowwise() |> mutate(d_fill_target=fill_es(target_ES_type, target_ES_value, target_N))|> 
  ggplot(aes(x=d_fill_target, y=target_d_calc))+geom_point()+geom_abline(slope=1,intercept=0)+geom_abline(slope=2,intercept=0, color="red")+coord_cartesian(xlim=c(0,5),ylim=c(0,5))

test |> rowwise() |> mutate(d_fill_rep=fill_es(replication_ES_type, replication_ES_value, replication_N))|>
  ggplot(aes(x=d_fill_rep, y=rep_d_calc))+geom_point()+geom_abline(slope=1,intercept=0)+geom_abline(slope=2,intercept=0, color="red")

@@

d_es <-  test |> 
  rowwise() |> 
  mutate(target_se_calc=est_SE(target_p_calc,target_d_calc),
         rep_se_calc=est_SE(rep_p_calc,rep_d_calc))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = F)
library(tidyverse)
library(viridis)
library(Replicate)
library(metafor)
library(esc)
theme_set(theme_bw())
```

# TODOs
- why are some ES's  about 2 times bigger than they should be? (from within v between stuff)
- validate df against num people?
 - what to do for chisq, other things? for ES

@@

  select(target_lastauthor_year, status, academic_year, subfield, target_year, stanford_internal, 
         open_data, open_materials, target_on_turk, replication_on_turk, 
         within_between, single_vignette, repeated_measure,
         target_N, target_test, target_raw_stat, target_pvalue, target_ES_type, target_ES_value,
         same_dir,
         replication_N, replication_test, replication_raw_stat, replication_pvalue, replication_ES_type, replication_ES_value,
         replicated_instructor_code, replicated_report_code)

@@

We parse out values from (for now) t tests, F tests, and chisq. 
Other stats to be added later, chisq ES to be added.

@@

parse_f <- function(fval, within_between){
  df_1 =str_extract(fval, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  df_2 =str_extract(fval, ",.*\\)") |> str_sub(2,-2) |> as.numeric()
  val=str_extract(fval, "=.*") |> str_sub(2,-1) |> as.numeric() |> abs()
  pval=pf(q=val, df1=df_1, df2=df_2, lower.tail=FALSE)
  d_calc=NA
  if(!is.na(df_1)&&df_1==1){
    if(within_between=="between"){
      d_calc=2*sqrt(val)/sqrt(df_2) # if there's just two groups, then F=t**2 and we can use the t-test thingy
    }
    else{
      d_calc=sqrt(val)/sqrt(df_2)
    }
  }
  else if(!is.na(df_1)){
    # we go through partial eta
    n=df_1+df_2+1 # rederive n b/c it's not always matching the n we have
    partial_eta=(val*df_1)/(val*df_1+df_2)
    d_calc=sqrt(n/(n-1)*(partial_eta/(1-partial_eta)))
  }
  return(data.frame("df_1"=df_1,"df_2"=df_2,"tstat"=NA, "fstat"=val, "p_calc"=pval, "d_calc"=d_calc))
}

parse_chisq <- function(chival, within_between){
  df=str_extract(chival, "\\(.*,") |> str_sub(2,-2) |> as.numeric()
  n=str_extract(chival, ",N=.*\\)") |> str_sub(4,-2) |> as.numeric()
  val=str_extract(chival, "\\)=.*") |> str_sub(3,-1) |> as.numeric() |> abs()
  print(val)
  pval=pchisq(val,df, lower.tail=F)
  phi=sqrt(val/n)
  d_calc=NA
  if (!is.na(df)&&df==1){
    d_calc=esc_phi(phi,pval,n, es.type="d")$es
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc))
}

#foo = "MCI m1=3.41[3.24,3.55], m2=3.44[3.24,3.57]"
parse_mean_ci <- function(mci,within_between,n){
  cond1=str_extract(mci, "m1.*\\],")
  cond2=str_extract(mci, "m2.*\\]")
  m1=str_extract(cond1, "m1=.*\\[") |> str_sub(4,-2) |> as.numeric()
  m2=str_extract(cond2, "m2=.*\\[") |> str_sub(4,-2) |> as.numeric()
  low1=str_extract(cond1, "\\[.*,.") |> str_sub(2,-3) |> as.numeric()
  low2=str_extract(cond2, "\\[.*,") |> str_sub(2,-2) |> as.numeric()
  high1=str_extract(cond1, ",.*\\]") |> str_sub(2,-2) |> as.numeric()
  high2=str_extract(cond2, ",.*\\]") |> str_sub(2,-2) |> as.numeric()
  se1=high1-low1/(2*1.96)
  se2=high2-low2/(2*1.96)
  se=(se1+se2)/2
  tval=(m1-m2)/se |> abs()
  if (within_between=="within"){
  df=n-1
  pval=pt(q=tval, df=df, lower.tail=FALSE)*2
  d_calc=tval/(sqrt(df)) 
  }
  else{
    df=NA
    pval=NA
    d_calc=NA
  }
  return(data.frame("df_1"=df,"df_2"=NA,"tstat"=tval, "fstat"=NA, "p_calc"=pval, "d_calc"=d_calc))
}

#parse_mean_ci(foo)

do_blanks <- function(){
  return(data.frame("df_1"=NA,"df_2"=NA,"tstat"=NA, "fstat"=NA, "p_calc"=NA, "d_calc"=NA))
}
do_parsing=function(raw_stat, within_between,n){
  if (is.na(raw_stat)) {return (do_blanks())}
  if (str_sub(raw_stat,1,1)=="t"){return(parse_t(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="F"){return(parse_f(raw_stat, within_between))}
  if (str_sub(raw_stat,1,1)=="c"){return(parse_chisq(raw_stat, within_between))}
  if (str_sub(raw_stat, 1,3)=="MCI"){return(parse_mean_ci(raw_stat, within_between,n))}
  return (do_blanks())
}
```

test <- d_val |> 
  mutate(target_raw_stat=gsub(" ","",target_raw_stat),
         target=pmap(list(target_raw_stat, within_between,target_N), do_parsing),
         replication_raw_stat=gsub(" ","",replication_raw_stat),
         rep=pmap(list(replication_raw_stat, within_between,replication_N), do_parsing)) |> 
  unnest(cols=c(target, rep), names_sep="_") |> 
  rowwise() |> 
  mutate(d_fill_target=fill_es(target_ES_type, target_ES_value, target_N),
         d_fill_rep=fill_es(replication_ES_type, replication_ES_value, replication_N))

@@

Comparing calculated p vals to provided (or hand calculated) ones. Line is x=y.

These look pretty reasonable. Lots on the line and others is places were <.001's happened.

@@

test |> 
  ggplot(aes(x=d_fill_target, y=target_d_calc))+geom_point()+geom_abline(slope=1,intercept=0)+geom_abline(slope=2,intercept=0, color="red")+coord_cartesian(xlim=c(0,5),ylim=c(0,5))+geom_abline(slope=.5,intercept=0, color="red")

test |> 
  ggplot(aes(x=d_fill_rep, y=rep_d_calc))+geom_point()+geom_abline(slope=1,intercept=0)+geom_abline(slope=2,intercept=0, color="red")+geom_abline(slope=.5,intercept=0, color="red")

## what to do with the weirdos

Where there's a difference, I could check whether it's being compared to something *I* calculated versus something that an original paper or students paper calculated? Except that there's sources for errors in any of those. Ones where we get an off  by two error -- should I double check within-betweeness for the test?

```{r}
test |> filter(target_d_calc>1.9*d_fill_target)
test |> filter(target_d_calc<.6*d_fill_target)

test |> filter(rep_d_calc>1.9*d_fill_rep)
test |> filter(rep_d_calc<.6*d_fill_rep)

## what didn't parse

test |> filter(is.na(target_d_calc)) |> filter(!is.na(target_raw_stat)) |>  pull(target_raw_stat)

test |> filter(is.na(rep_d_calc)) |> filter(!is.na(replication_raw_stat)) |> pull(replication_raw_stat)
```

# Where does df not line up?
Note: it's not *necessarily* a problem if df's don't line up -- there could be some repeated measures or welch's test stuff going on, but nice to check.

test

@@

d_es <-  test |> 
  rowwise() |> 
  mutate(target_se_calc=est_SE(target_p_calc,target_d_calc),
         rep_d_calc=ifelse(same_dir=="no", -rep_d_calc, rep_d_calc),
         rep_se_calc=est_SE(rep_p_calc,rep_d_calc))
